---
title: "Data cleaning"
author: "Tobias Schwoerer"
date: "November 28, 2017"
output: html_document
---
Go to the Knb, copy the URL for the data I want 
```{r}
catch <- read.csv("df35b.302.1.csv", stringsAsFactors = FALSE)
```
On a mac I could use the URL in quotes but on a PC that doesn't work as smoothly. But this following code didn't really work. Try https://stackoverflow.com/questions/4125788/reading-csv-files-over-ssl-with-r
```{r}
require(RCurl)
myCsv <- getURL("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", stringsAsFactors = FALSE)
temporaryFile <- tempfile()
con <- file(temporaryFile, open = "w")
cat(myCsv, file = con) 
close(con)

read.csv(temporaryFile)
```

```{r}
library(dplyr)
library(tidyr)
```
Let's do some manipulations, here we delete the columns we don't need called All and notesRegCode
```{r}
catch <-catch %>% 
  select(-All, -notesRegCode)
```
Now we switch the wide formatted data into a tall dataset using the gather function, and create a new dataset that we also call catch and basicaly overwrite the old datafile

```{r}
catch <- catch %>%
  gather(species, catch, -Region, - Year)
```


```{r}
summary(catch)
catch_integers <- as.integer(catch$catch)
# figuring out which ones are NA
nas <- which(is.na(catch_integers))
# The ADFG report shows that an i was entered when it should have been a 1
catch[nas,"catch"] <- 1
catch[401,]
summary(catch)
```
The summary output still shows that we have some characters in the catch variable
```{r}
catch <- catch %>%
  mutate(catch = as.integer(catch))
summary(catch)
```
Looking at metadata on KNB showed that catch is in thousands of fish, so we need to multiply by 1000 on catch variable
```{r}
catch <- catch %>%
  mutate(catch = catch*1000)
summary(catch)
```

```{r}
catch %>%
  group_by(Region) %>%
  summarize(mean_catch=mean(catch))
```

```{r}
sample_size <- catch %>%
  group_by(Region) %>%
  summarize(n=n())
summary(sample_size)
```
Filtering rows, say SOuthern southeast catches

```{r}
sse_catches <- catch %>%
  filter(Region =="SSE")
sse_catches
```
Filtering really small catches in Southern Southeast
```{r}
small_catches <- catch %>%
  filter(catch <1000000)
summary(small_catches)
small_catches
```
Combined filtering on multiple conditions
Large sse catches
```{r}
large_sse_catch <- catch %>%
  filter(1000000 & Region == "SSE")
summary (large_sse_catch)
large_sse_catch
#Doesn't seem to be sorted, so sorting it
large_sse_catch %>%
  arrange(species, Year)
```
Joining two datasets together

```{r}
region_defs <- read.csv("df35b.303.1.csv", stringsAsFactors = FALSE)
View(region_defs)
View(catch)
```



```{r}
catch %>%
  group_by(Region) %>%
  summarise(total_catch = sum(catch)) %>%
  left_join(region_defs, by = c("Region" = "code"))
```
Example for Separating dates that are fortunately all formatted equally 

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)
dates_df
dates_df <- dates_df %>%
  separate(date, c("month","day","year"),"/")
dates_df
#undo what we just did creating a new variable called date
dates_df %>%
  unite(date, month, day, year, sep = "/")
```

