---
title: "My website"
author: "Tobias Schwoerer"
date: "November 29, 2017"
output: html_document
---
Use echo=FALSE to leave all the code cunchs out of the website
the cache=TRUE option leaves output from the first dun in the cache speeding up any additional rendering that is necessary.

```{r, echo=FALSE}
escapement <- read.csv("knb.92020.1.csv",stringsAsFactors = FALSE)

```

```{r}
library(dplyr)
library(tidyr)
library(leaflet)
library(ggplot2)
library(DT)
```

To calculate the median annual escapement by species, we first need to separate the date column to get a year column, we do this by also creating month and day columns. We use the separate function to do this. 

```{r}
median_esc <- escapement %>%
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-")%>%
  group_by(Species, Location, Year) %>%
  summarise(escapement = sum(DailyCount, na.rm=TRUE))  %>%
  group_by(Species) %>%
  summarise(median_escapement = median(escapement))
median_esc
```
See above, the best way to sum is to use the na.rm argument within the sum function  which ignores NA in calculating the sum. 

```{r}
ggplot(median_esc, aes(Species, median_escapement)) +
  geom_col() +
  coord_flip()

```

```{r}
datatable(median_esc)
```

How to use the leaflet package making an interactive map

```{r}
locations <- escapement %>%
  distinct(Location, Latitude, Longitude) %>%
  drop_na()

leaflet(locations) %>%
  addTiles() %>%
  addMarkers(~ - abs(Longitude), ~ Latitude, popup = ~ Location)

```